---
title: "CRPS profile categorization"
output: 
  flexdashboard::flex_dashboard:
    css: my-custom-rules.css
    theme: lumen
    source_code: null
runtime: shiny
---

<style>
.page-footer {
  display: flex;
  justify-content: center;
  align-items: center;
  color: lightgrey;
  font-size: 10px;
  padding: 10px 15px;
  margin-top: 25px; /* push it away from content */
  flex-wrap: wrap;
  text-align: center;
}

.page-footer div {
  flex: 1 1 35%;
}

@media (max-width: 600px) {
  .page-footer {
    flex-direction: column;
    align-items: center;
    text-align: center;
    font-size: 10px;
  }

  .page-footer div {
    flex: unset;
    margin: 5px 0;
  }
}

</style>



```{r library}
library(shiny)
library(tidyverse)
library(readxl)
library(RColorBrewer)
library(ggthemes)
library(patchwork)
library(dplyr)
library(ggtext)
library(glue)
library(grid)
library(gridExtra)
library(DT)
library(here)
library(plotly)



```

```{r data}
#data_ind <- read.csv(here("data", "data_new_pt.csv"), sep = ";") 


norm <- read_xlsx("norms_sdrc_20240307.xlsx") # importing normative values


data_norm <- norm |> pivot_longer(cols = starts_with(c("h",  "f")),
                                  names_to = c("sex", "metrics", "age_cat"),
                                  names_sep = "_",
                                  values_to = c("value")) |> 
  pivot_wider(names_from = "metrics",
              values_from = "value") |> 
  mutate(age_cat = factor(age_cat)) |>  rename("qst_modality" = "test_mod", "area_qst" = "site") |> # normative values areas are categorized according to extremity
  mutate(extremity = if_else(area_qst == "foot", "LE", "UE")) |> 
  select(!area_qst)



```


```{r functions}
## Scoring function to be called per record
lg_scoring<-function(dat) {
  # Create auxiliary variables 
  if(is.na(dat$css_signs___2)) {
    css_signs___2_lg<-0
    css_signs___2_lg_m<-1
  }
  else {
    css_signs___2_lg<-dat$css_signs___2
    css_signs___2_lg_m<-0
  }
  if(is.na(dat$orebro_score)) {
    orebro_score_lg<-0
    orebro_score_lg_m<-1
  }
  else {
    orebro_score_lg<-dat$orebro_score
    orebro_score_lg_m<-0
  }
  if(is.na(dat$rbcrpsbpds_score)) {
    rbcrpsbpds_score_lg<-0
    rbcrpsbpds_score_lg_m<-1
  }
  else {
    rbcrpsbpds_score_lg<-dat$rbcrpsbpds_score
    rbcrpsbpds_score_lg_m<-0
  }
  if(is.na(dat$triggering_event_no0_no1)) {
    triggering_event_no0_no1_lg<-0
    triggering_event_no0_no1_lg_m<-1
  }
  else {
    triggering_event_no0_no1_lg<-dat$triggering_event_no0_no1
    triggering_event_no0_no1_lg_m<-0
  }
  if(is.na(dat$ppt_z_norm)) {
    ppt_z_norm_lg<-0
    ppt_z_norm_lg_m<-1
  }
  else {
    ppt_z_norm_lg<-dat$ppt_z_norm
    ppt_z_norm_lg_m<-0
  }

  # Compute classification logits 
  Cluster_lg_1<-(4.5188594)+
     (-0.59396254)*css_signs___2_lg+
     (-0.16227453)*orebro_score_lg+
     (0.4033855)*rbcrpsbpds_score_lg+
     (0.74837457)*triggering_event_no0_no1_lg+
     (0.277063)*ppt_z_norm_lg+
     (0.00085489678)*orebro_score_lg*orebro_score_lg+
     (-0.010732156)*rbcrpsbpds_score_lg*rbcrpsbpds_score_lg+
     (-0.11925119)*ppt_z_norm_lg*ppt_z_norm_lg+
     (-0.42008314)*css_signs___2_lg_m+
     (-6.6429874)*orebro_score_lg_m+
     (1.2273978)*rbcrpsbpds_score_lg_m+
     (2.1113198)*triggering_event_no0_no1_lg_m+
     (-0.36593027)*ppt_z_norm_lg_m
  Cluster_lg_2<-(18.784051)+
     (-1.4680407)*css_signs___2_lg+
     (-0.2917661)*orebro_score_lg+
     (-0.3933944)*rbcrpsbpds_score_lg+
     (0.45623914)*triggering_event_no0_no1_lg+
     (-0.20779313)*ppt_z_norm_lg+
     (0.0016569632)*orebro_score_lg*orebro_score_lg+
     (-0.019228013)*rbcrpsbpds_score_lg*rbcrpsbpds_score_lg+
     (-0.054071223)*ppt_z_norm_lg*ppt_z_norm_lg+
     (-0.68517065)*css_signs___2_lg_m+
     (-10.360489)*orebro_score_lg_m+
     (-8.1253055)*rbcrpsbpds_score_lg_m+
     (1.1429389)*triggering_event_no0_no1_lg_m+
     (-0.67164784)*ppt_z_norm_lg_m
  Cluster_lg_3<-(-25.091576)+
     (1.2892414)*css_signs___2_lg+
     (0.35587142)*orebro_score_lg+
     (0.8039539)*rbcrpsbpds_score_lg+
     (-0.066881621)*triggering_event_no0_no1_lg+
     (-0.037514539)*ppt_z_norm_lg+
     (-0.0026335818)*orebro_score_lg*orebro_score_lg+
     (0.00021162458)*rbcrpsbpds_score_lg*rbcrpsbpds_score_lg+
     (0.05038911)*ppt_z_norm_lg*ppt_z_norm_lg+
     (0.75162977)*css_signs___2_lg_m+
     (10.887568)*orebro_score_lg_m+
     (13.831902)*rbcrpsbpds_score_lg_m+
     (-0.42588956)*triggering_event_no0_no1_lg_m+
     (-0.016862815)*ppt_z_norm_lg_m
  Cluster_lg_4<-(1.7886653)+
     (0.77276184)*css_signs___2_lg+
     (0.098169212)*orebro_score_lg+
     (-0.813945)*rbcrpsbpds_score_lg+
     (-1.1377321)*triggering_event_no0_no1_lg+
     (-0.031755329)*ppt_z_norm_lg+
     (0.00012172176)*orebro_score_lg*orebro_score_lg+
     (0.029748545)*rbcrpsbpds_score_lg*rbcrpsbpds_score_lg+
     (0.1229333)*ppt_z_norm_lg*ppt_z_norm_lg+
     (0.35362402)*css_signs___2_lg_m+
     (6.1159092)*orebro_score_lg_m+
     (-6.9339944)*rbcrpsbpds_score_lg_m+
     (-2.8283691)*triggering_event_no0_no1_lg_m+
     (1.0544409)*ppt_z_norm_lg_m

  # Compute odds from logits 
  max_lg<-Cluster_lg_1
  if(Cluster_lg_2>max_lg) {
    max_lg<-Cluster_lg_2
  }
  if(Cluster_lg_3>max_lg) {
    max_lg<-Cluster_lg_3
  }
  if(Cluster_lg_4>max_lg) {
    max_lg<-Cluster_lg_4
  }
  Cluster_lg_1<-exp(Cluster_lg_1-max_lg)
  Cluster_lg_2<-exp(Cluster_lg_2-max_lg)
  Cluster_lg_3<-exp(Cluster_lg_3-max_lg)
  Cluster_lg_4<-exp(Cluster_lg_4-max_lg)

  # Compute modal class and probabilities from odds 
  max_lg<-Cluster_lg_1
  Cluster_lg_modal<-1
  if(Cluster_lg_2>max_lg) {
    max_lg<-Cluster_lg_2
    Cluster_lg_modal<-2
  }
  if(Cluster_lg_3>max_lg) {
    max_lg<-Cluster_lg_3
    Cluster_lg_modal<-3
  }
  if(Cluster_lg_4>max_lg) {
    max_lg<-Cluster_lg_4
    Cluster_lg_modal<-4
  }
  sum_lg<-Cluster_lg_1+Cluster_lg_2+Cluster_lg_3+Cluster_lg_4
  Cluster_lg_1<-Cluster_lg_1/sum_lg
  Cluster_lg_2<-Cluster_lg_2/sum_lg
  Cluster_lg_3<-Cluster_lg_3/sum_lg
  Cluster_lg_4<-Cluster_lg_4/sum_lg

  return(list(
    "Cluster_modal"=Cluster_lg_modal,
    "Cluster_1"=Cluster_lg_1,
    "Cluster_2"=Cluster_lg_2,
    "Cluster_3"=Cluster_lg_3,
    "Cluster_4"=Cluster_lg_4
  ))
}
```

# Description

**Description**

This [shiny](https://shiny.posit.co/r/getstarted/shiny-basics/lesson1/) application provides a probability to belong to certain *early CRPS profiles* (< 6 months). These profiles have been associated with different clinical evolutions and are based on the results of a latent class analysis (LCA), using five variables, considered to be involved in several pathophysiological mechanisms (see [4] for further explanation):
<br>


- The presence of *allodynia* to light touch, cold, light pressure or gentle joint mobilisation (sign) [2].
- The physical intensity of the *triggering event* [1], classified as (1) no or minor incident/soft tissue injury, (2) minor surgical procedure (i.e., elective surgery), (3) fracture not requiring surgical repair, (4) major injury requiring surgical repair.
- The *Örebro Musculoskeletal Pain Screening questionnaire - short form* ([ÖMPSQ-SF](https://orthotoolkit.com/ompsq-sf/)), which assesses psychosocial severity [3]. 
- Deep tissue sensitivity, using the standardized *pressure pain threshold* (PPT) [5].
- The *revised-Bath-CRPS-Body perception disorder scale* ([r-B-CRPS-BPDS](https://www.sciencedirect.com/science/article/pii/S1526590021002157#sec0032)) [7]. 

The application combines patient information (age, sex, affected extremity) with individual PPT value (kPa) to compute a Z-score using the German normative values [5] with the following formula:$$\text{PPT Z-score} = \frac{\text{PPT}_{\text{individual}} - \text{Mean}_{\text{database}}}{\text{SD}_{\text{database}}}$$


PPT should be obtained using the DFNS protocol (average of three measurements, ideally on the back of the hand or foot to match normative data, but in the CRPS field, on the most painful area [6]).


No data is stored and the application is free to use.

**Contact**


[Marc-Henri Louis](marc-henri.louis@uclouvain.be), Institute of Neuroscience, UCLouvain, Belgium.

**References**


- [1] 	Bean DJ, Johnsond MH, Heiss-Dunlop W, Kydd RR. Extent of recovery in the first 12 months of complex regional pain syndrome type-1: A prospective study. European Journal of Pain 2016;20:884–894.

- [2] 	Goebel A, Birklein F, Brunner F, Clark JD, Gierthmühlen J, Harden N, Huygen F, Knudsen L, McCabe C, Lewis J, Maihöfner C, Magerl W, Moseley GL, Terkelsen A, Thomassen I, Bruehl S. The Valencia consensus-based adaptation of the IASP complex regional pain syndrome diagnostic criteria. Pain 2021;162:2346–2348.

- [3] 	Linton SJ, Nicholas M, MacDonald S. Development of a short form of the Örebro Musculoskeletal Pain Screening Questionnaire. Spine (Phila Pa 1976) 2011;36:1891–1895.

- [4] 	Louis M-H, Legrain V, Aron V, Filbrich L, Henrard S, Barbier O, Libouton X, Mouraux D, Lambert J, Berquin A. Early CRPS Is a Heterogeneous Condition: Results From a Latent Class Analysis. Eur J Pain 2025;29:e4785.

- [5] 	Magerl W, Krumova EK, Baron R, Tölle T, Treede R-D, Maier C. Reference data for quantitative sensory testing (QST): refined stratification for age and a novel method for statistical comparison of group data. Pain 2010;151:598–605.

- [6] 	Rolke R, Baron R, Maier C, Tölle TR, Treede -D. R., Beyer A, Binder A, Birbaumer N, Birklein F, Bötefür IC, Braune S, Flor H, Huge V, Klug R, Landwehrmeyer GB, Magerl W, Maihöfner C, Rolko C, Schaub C, Scherens A, Sprenger T, Valet M, Wasserka B. Quantitative sensory testing in the German Research Network on Neuropathic Pain (DFNS): standardized protocol and reference values. Pain 2006;123:231–243.

- [7] 	Ten Brink AF, Halicka M, Vittersø AD, Jones HG, Stanton TR, Bultitude JH. Validation of the Bath CRPS Body Perception Disturbance Scale. J Pain 2021. 






<div class="page-footer">
  <div>Application designed by [Vladimir Aron](https://github.com/vladaron/CRPS_longitudinal) (Institute of Neuroscience, UCLouvain) using R packages  `shiny` and `flexdashboard`..</div>
</div>

# Application

## {.sidebar data-width=350}

###

**Patient information**

```{r inputs1}



radioButtons("sex", "Sex: ", choices = c("Male" = "h","Female" = "f"), inline = T)

sliderInput("age", "Age (years):", min = 18, max = 110, step = 1, value = 35)

radioButtons("extremity", "Affected extremity: ", choices = c("Upper extremity" = "UE","Lower extremity" = "LE"), inline = T)







```

###


**Profile variables**

```{r inputs2}


radioButtons("css", "Presence of allodynia: ", choices = c("Yes" = 1,"No" = 0), inline = T)

selectInput("trig", "Triggering event: ", choices = c("No or minor incident/soft tissue injury" = 1, "Minor surgical procedure" = 2, "Fracture not requiring surgical repair" = 3, "Major injury requiring surgical repair" = 4))

sliderInput("oreb", "ÖMPSQ - SF score:", min = 0, max = 100, step = 1, value = 10)

sliderInput("body", "r-Bath-CRPS-BPDS score:", min = 0, max = 47, step = 1, value = 10)


#sliderInput("ppt", "PPT score (kPa):", min = 1, max = 1500, step = 1, value = 150)

numericInput("ppt", "PPT score (kPa):", min = 1, max = 1500, step = 1, value = 150)

#sliderInput("z", "PPT z-score:", min = -3, max = 3, step = 0.01, value = 1)




```



```{r reactive}



age_cat <- reactive({
  age <- input$age
 # age <- 35
  age_cat_val <- case_when(
    age >= 18 & age < 30 ~ "1",
    age >= 30 & age < 40 ~ "2",
    age >= 40 & age < 50 ~ "3",
    age >= 50 & age < 60 ~ "4",
    age >= 60           ~ "5",
    TRUE                ~ NA_character_  # Handle unexpected cases
  )
  return(age_cat_val)

})



z_score <- reactive({
  #sex <- "h"
 # age_cat <- age_cat_val
 # extremity <- "UE"
  norm <- data_norm |> 
    filter(sex == input$sex, age_cat == age_cat() ,extremity == input$extremity, qst_modality == "ppt") 
  
   # norm <- data_norm |> 
   # filter(sex == "h", age_cat == age_cat_val ,extremity == "UE", qst_modality == "ppt") 
 # ppt <- 100
 #z_score <- (ppt- norm$m)/norm$sd
  z_score <- (log10(input$ppt) - norm$m)/norm$sd
  return(z_score)
})



#trig <- reactive({
 #  t <- input$trig
 
 # t_cat <- case_when(
  #  t == "Soft tissue injury/minimal trauma" ~ "1",
  #   t=="Elective surgery" ~ "2",
  #   t=="Trauma" ~ "3",
  #   t=="Trauma with fracture/major trauma" ~ "4"
#  )
#  trig <- as.numeric(t_cat)
#  return(trig)
  

#})

outdat <- reactive({
  


  df <- data.frame(css_signs___2 = input$css, orebro_score= input$oreb, rbcrpsbpds_score = input$body, triggering_event_no0_no1 = input$trig, ppt_z_norm = z_score() ) |> 
    mutate(across(everything(), as.numeric))
  
   # df <- data.frame(css_signs___2 = 2, orebro_score= 13, rbcrpsbpds_score = 4, triggering_event_no0_no1 = 1, ppt_z_norm = z_score)
  
  scoring<-lg_scoring(df[1,])
  outdat <- data.frame(
    Cluster_modal = scoring$Cluster_modal,
    "cluster_1" = 100*scoring$Cluster_1,
    "cluster_2" = 100*scoring$Cluster_2,
    "cluster_3" = 100*scoring$Cluster_3,
    "cluster_4" = 100*scoring$Cluster_4
  )
  outdat$z = z_score()
  return(outdat)
})
```



## 

### Probability of profile & PPT Z-score {data-height=100}

```{r table_out}



renderTable({
  outdat() |> 
    select(-Cluster_modal) |> 
    rename("Z score PPT" = z) |> 
    rename("Moderate CRPS" = "cluster_1") |> 
    rename("Mild CRPS" = "cluster_2") |> 
      rename("Body representation disturbance CRPS" = "cluster_3") |> 
    rename("Pressure allodynia CRPS" = "cluster_4") |> 
    select("Mild CRPS", "Moderate CRPS", "Body representation disturbance CRPS","Pressure allodynia CRPS", "Z score PPT" )
})


```


### {.no-mobile}

```{r}

renderPlotly({
 p <-  outdat() |> 
    select(-Cluster_modal, -z) |> 
    pivot_longer(cols = everything(), names_to = "cat", values_to = "proba") |> 
    mutate(cat = factor(cat, levels = c("cluster_2", "cluster_1", "cluster_3", "cluster_4"), labels = c("Mild CRPS", "Moderate CRPS", "Body representation disturbance CRPS","Pressure allodynia CRPS" ))) |> 
    ggplot(aes(y = proba, x = cat, fill = cat)) +
      geom_col() +
    scale_x_discrete(labels = NULL) +
    scale_y_continuous(limits = c(0,100), breaks = c(0, 25, 50,75,100)) +
    scale_fill_manual(values = c( "#288B00", "#283F8B","#F67E00", "#F43032")) +
    labs(fill = "", x = "", y = "%") + 
    theme_minimal() #+
   # theme(legend.position = "bottom",
          #legend.text = element_text(size = 11, colour = "black", family = "Roboto"))
 plotly::ggplotly(p) |> 
   plotly::layout(legend=list(x=0, 
                                 xanchor='left',
                                 yanchor='bottom',
                                 orientation='h')) 
})

#renderPlot({
  
# df <- data.frame(ppt = "PPT", val = z_score()) 
 #df|> 
   # ggplot(aes(x = ppt, y = val)) +
   # geom_point(size = 3, col = "red") +
   # scale_y_continuous(limits = c(-3,3)) +
   # geom_hline(yintercept = 0, col = "black", linetype = "dashed") +
   # labs(title = "Individual z score")
#})

```

### {.mobile}

```{r}

renderPlot({
 p <-  outdat() |> 
    select(-Cluster_modal, -z) |> 
    pivot_longer(cols = everything(), names_to = "cat", values_to = "proba") |> 
    mutate(cat = factor(cat, levels = c("cluster_2", "cluster_1", "cluster_3", "cluster_4"), labels = c("Mild CRPS", "Moderate CRPS", "Body representation disturbance CRPS","Pressure allodynia CRPS" ))) |> 
    ggplot(aes(y = proba, x = cat, fill = cat)) +
      geom_col() +
    scale_x_discrete(labels = NULL) +
    scale_y_continuous(limits = c(0,100), breaks = c(0, 25, 50,75,100)) +
    scale_fill_manual(values = c( "#288B00", "#283F8B","#F67E00", "#F43032")) +
    labs(fill = "", x = "", y = "%") + 
    theme_minimal() +
   theme(
     legend.position = "bottom",
     legend.text = element_text(size = 7, family = "Roboto"),
     plot.margin = margin(10, 10, 30, 10)
    # legend.key.width = unit(0.5, "cm")
   ) + 
   guides(fill = guide_legend(nrow = 2, byrow = TRUE))

 p

})


```



### Interpretation {data-height=100}

```{r txt}
renderText({
  
  cat <- switch (outdat()$Cluster_modal,
          "1" = "Moderate CRPS",
          "2" = "Mild CRPS",
          "3" = "Body representation disturbance CRPS",
          "4" = "Pressure allodynia CRPS"
  )
  
  
  df_sort <- outdat() |> 
    select(-Cluster_modal, -z) |> 
  pivot_longer(cols = everything(), names_to = "cat", values_to = "proba")
  
  max <- round(max(df_sort$proba),2)
  
  if(max >= 80){
    out <- paste("Your patient can be classified as", cat, "with a probability of ", max,"%.") 
  }else{
     out <- paste(
       "Your patient can be classified as", cat, "with a probability of ", max, "%.",
       "Please note that a criteria of 80% should be used to attribute a profile. Here, this criteria is not reached. Results should be interpreted as a tendency towards a profile rather than a definite categorization.")
  }
  return(out)
})
```









